---
title: "Social Network Analysis <br/> Home Assignment 3"
author: 'Shupeyko Nikita'
date: "due date - 05.06.2016 23:59"
output:
  pdf_document:
    latex_engine: xelatex
    toc: yes
  html_document:
    toc: yes
---
# Network communities

```{r, message = F, echo = F, warning = F}

library(igraph)
library(xtable)
options(xtable.comment = F)

```

## Yahoo Music Network

For this assignment, you have to load [the part of Yahoo Music Network](https://drive.google.com/file/d/0Bwd5a7NScgMtNElfOFdSbE1UYnM/view). Archive contains network in GML format and corresponding list of artists.

```{r}

yg.f = "music2K.gml"
yg = read.graph(yg.f, format = c("gml"))

```

Edges in this network appear if enough number of users have given ratings to both music bands. Note, that edges are weighted with similarity of the ratings.



### 1. Preprocessing

This network needs to be preprocessed. Do the steps, described on seminar, to make it look properly. How many vertices and edges were deleted?

```{r, results = "asis", fig.width = 10, fig.height = 10}

# Deleting isolated nodes.
yg.isolated.vertices= V(yg)[degree(yg) == 0]
yg.delete.vertices.diff = vcount(yg)
yg = delete.vertices(yg, degree(yg) == 0)
yg.delete.vertices.diff = yg.delete.vertices.diff - vcount(yg)
paste("Deleted vertex count: ", yg.delete.vertices.diff)

# Simplifying the graph's overall structure
# (get rid of loops and multiedges).
yg.deleted.edge.diff = ecount(yg)
yg = simplify(yg)
yg.deleted.edge.diff = yg.deleted.edge.diff - ecount(yg)
paste("Deleted edge count: ", yg.deleted.edge.diff)

# Obtaining descriptives.
func_print.graph.descriptives = function(g) {
  g.n = vcount(g)
  g.m = ecount(g)
  g.dens = graph.density(g)
  g.avgdeg = 2 * g.m / g.n
  g.avgpathlen = average.path.length(g)
  g.diam = diameter(g)
  tbl = rbind(g.n,
              g.m,
              g.dens,
              g.avgdeg,
              g.avgpathlen,
              g.diam)
  rownames(tbl) = c("Node count",
                    "Edge count",
                    "Network Density",
                    "Average Node Degree",
                    "Average Path Length",
                    "Diameter")
  colnames(tbl) = c("Value")
  print(xtable(tbl),
        type = "latex",
        floating = T, 
        include.rownames = T,
        latex.environments = "center")
}
func_print.graph.descriptives(yg)


func_plot.graph.shortestpathsdistribution = function(g) {
  g.plh = path.length.hist(g)$res
  g.plh.tbl = as.table(g.plh)
  names(g.plh.tbl) = 1:length(g.plh.tbl)
  g.plh.bp = barplot(g.plh.tbl, 
                     main = "Shortest Path Length Distribution",
                     xlab = "Path length",
                     ylab = "Paths total")
  text(x = g.plh.bp, 
       y = g.plh, 
       label = g.plh, 
       cex = 0.8, 
       pos = 3)
}
func_plot.graph.shortestpathsdistribution(yg)

```



### 2. Clustering

Define clusters for this networks using one of the algorithms described on lectures and seminars. Compute density of obtained clusters:

```{r, results = "asis"}

# Thanks to its speed on large graphs, the Louvain algorithm is used 
# to reveal communities in the given network.
yg.communities = multilevel.community(yg)
yg.communities.count = length(groups(yg.communities))

# yg.communities.as.subgraphs = c()
yg.communities.as.subgraphs = vector("list", yg.communities.count)
for (i in seq_along(yg.communities.as.subgraphs)) { 
    yg.communities.as.subgraphs[[i]] = induced.subgraph(yg, yg.communities$membership == i)
}

yg.community.densities = vector("list", yg.communities.count)
for (i in seq_along(yg.community.densities)) {
  yg.community.densities[[i]] = graph.density(yg.communities.as.subgraphs[[i]])
}
yg.community.names = vector("list", yg.communities.count)
for (i in seq_along(yg.community.names)) {
  yg.community.names[[i]] = paste("Community #", i)
}

tbl = cbind(as.matrix(yg.community.names),
            as.matrix(yg.community.densities))
colnames(tbl) = c("", "Density")
tbl = xtable(tbl)
align(tbl) = "c|r|c"
print(tbl, 
      floating = T, 
      type = "latex",
      include.rownames = F,
      latex.environments = "center")

```

Compute the ratio of inner clusters connections to outer ones:

```{r, results = "asis"}

yg.m = ecount(yg)
yg.community.inoutratios = vector("list", yg.communities.count)
for (i in seq_along(yg.community.inoutratios)) {
  c = yg.communities.as.subgraphs[[i]]
  c.m = ecount(c)
  yg.community.inoutratios[[i]] = c.m / ( yg.m - c.m )
}

tbl = cbind(as.matrix(yg.community.names),
            as.matrix(yg.community.inoutratios))
colnames(tbl) = c("", "Inner-Outer Connections Ratio")
tbl = xtable(tbl)
digits(tbl) = c(0, 0, 4)
align(tbl) = "c|r|c"
print(tbl, 
      floating = T, 
      type = "latex",
      include.rownames = F,
      latex.environments = "center")

```



### 3. Visualization & interpretation

Visulize five of the most dense clusters. Use names of artists as node labels on the graph.

```{r, fig.width = 10, fig.height = 10}

yg.artists = scan("artists.txt", what = character(), sep = "\n")
yg.artists = yg.artists[-yg.isolated.vertices]
V(yg)$label = yg.artists

func_top.N.indecies = function(l, N) {
  l = unlist(l)
  N = as.integer(N)
  which( l > sort(l)[ length(l)-N ] )
}

yg.communities.top.5.density.indecies = func_top.N.indecies(yg.community.densities, 5)
yg.communities.top.5.density = yg.communities[yg.communities.top.5.density.indecies]
yg.vertices.in.communities.top.5.density = V(yg)[yg.communities$membership %in% yg.communities.top.5.density.indecies] 

# palette()
crp = colorRampPalette(c("red", "blue", "yellow", "pink", "maroon"))(5)
for (v.i in yg.vertices.in.communities.top.5.density) {
  V(yg)[v.i]$color = crp[which( yg.communities.top.5.density.indecies == yg.communities$membership[v.i] )]
  V(yg)[v.i]$label = yg.artists[v.i]
}

yg.subgraph.vertices.in.communities.top.5.density = induced.subgraph(yg, yg.vertices.in.communities.top.5.density)

# V(yg.subgraph.vertices.in.communities.top.5.density)$color = 7
plot(yg.subgraph.vertices.in.communities.top.5.density)

```

<b> (Extra task) </b>Try to interpret (using Wikipedia or other resources) obtained clusters.

Artist clusters colored by genre:

*Sky Blue* --- **Pop/Folk**;

*Blue* --- **Jazz/Blues**;

*Dark Green*--- **Gospel**;

*Orange* --- **Electro/Hip-hop/Dance**

*Yellow* --- **Rock/Punk-rock**
