---
title: "Social Network Analysis <br/> Home Assignment 4"
author: 'Shupeyko Nikita'
date: "due date - 12.06.2016 23:59"
output:
  pdf_document:
    toc: yes
  html_document:
    toc: yes
---
# Network Epidemics

```{r, echo = F, message = F, warning = F}

library(igraph)
library(EpiModel)
library(xtable)
options(xtable.comment = F)

```

## SIR Model

You need to perform epidemic SIR model on different types of networks:
Try different parameters for network generation: 

```{r}

gs = list()

# The number of vertices must be the same among the graphs,
# and so are edge counts/edge probabilities.
# The graph represented by Net.txt is the largest one.
gs$ptp = read.graph("Net.txt")
gs$ptp = as.undirected(gs$ptp)
gs$ptp$name = "Net"
ptp.n = vcount(gs$ptp)
ptp.m = ecount(gs$ptp)
ptp.avg.node.deg = 2*ptp.m / ptp.n
ptp.p = 2*ptp.m / (ptp.n*(ptp.n-1))

gs$ba = barabasi.game(n = ptp.n, 
                      m = ptp.avg.node.deg,
                      directed = F)
gs$ba$name = "Barabasi"
 
gs$er = erdos.renyi.game(n = ptp.n, 
                         p.or.m = ptp.m,
                         type = c("gnm"), 
                         directed = F)
gs$er$name = "Erdos-Renyi"

gs$ws = watts.strogatz.game(dim = 1, 
                            size = ptp.n,
                            nei = 3, 
                            p = ptp.p)
gs$ws$name = "Watts-Strogatz"

``` 

Moreover, perform modeling on real peer-to-peer network [here](https://www.hse.ru/data/2016/04/21/1130159900/Net.txt)

Your goal is to perform a research on epidemics:

```{r, warning = F}

simulate.sir.on.graphs = function(graphs,
                                  beta,
                                  gamma,
                                  no.sim) {
  sim = list(results = lapply(graphs,
                              sir,
                              beta = beta, 
                              gamma = gamma, 
                              no.sim = no.sim),
             params =  list(beta = beta, 
                            gamma = gamma, 
                            no.sim = no.sim))
  return(sim)
}

# For the sake of consistency, the same number of trials 
# was kept across all simulations.
sims = list(`1` = simulate.sir.on.graphs(gs, 3, 9, 250),
            `2` = simulate.sir.on.graphs(gs, 6, 6, 250),
            `3` = simulate.sir.on.graphs(gs, 9, 3, 250))

```

at least 3 different versions, for example:

* betta (4 6 8)
* gamma (8 6 2)
* niter (100 500 1000)
    
*For some reason `beta` and `gamma` parameters should not be set below 0 and 1. Looks like they are somehow normilized during simulation.*

You need to plot three values on the graphics: Number of infected, number of suseprible, number of recovered - all depends on time.
As a result of this task, you need to provide 12 plots (one for each network with 3 diferent parameters) with explanation.`

```{r}

plot.simulation = function(simulation,
                           graph.colors,
                           kind = c("NS", "NI", "NR")) {
  if (length(simulation) == 0) {
    stop("simulation is NULL or empty!")
  }
  if (length(simulation$results) == 0) {
    stop("simulation$results is NULL or empty!")
  }
  if (length(simulation$params) == 0) {
    stop("simulation$params is NULL or empty!")
  }
  graph.names = names(simulation$results)
  if (length(graph.names) != length(graph.colors)) {
    stop("The number of graphs in simulation does not match the number of colors specified!")
  }
  if (!(kind %in% c("NS", "NI", "NR"))) {
    stop("Cannot plot anything of kind other than 'NS', 'NI', or 'NR'!")
  }
  
  x.max = max(sapply(sapply(simulation$results, 
                            time_bins),
                     max))
  y.max = 1.05 * max(sapply(sapply(simulation$results, 
                                   function(x) median(x)[[kind]]), 
                            max, 
                            na.rm = T))
  for (graph.name.i in seq_along(graph.names)) {
    graph.name = graph.names[graph.name.i]
    results.for.graph = simulation$results[[graph.name]]
    if (graph.name.i == 1) {
      plot(time_bins(results.for.graph),
           median(results.for.graph)[[kind]], 
           type = "l",
           xlim = c(0, x.max), 
           ylim = c(0, y.max), 
           xlab = "Time", 
           ylab = paste0(kind, " (t)"),
           lwd = 2, 
           col = graph.colors[graph.name.i])
      title(kind)
      mtext(sprintf("beta = %i, gamma = %i, no.sim = %i",
                    simulation$params$beta,
                    simulation$params$gamma,
                    simulation$params$no.sim))
    } else {
      lines(time_bins(results.for.graph), 
            median(results.for.graph)[[kind]], 
            lwd = 2,
            col = graph.colors[graph.name.i])
    }
  }
  legend("topright", 
         graph.names,
         col = graph.colors, 
         lty = 1)
}


graph.colors = c("blue", "green", "orange", "red")

```



\newpage
### NS Comparison Charts

```{r, fig.width = 10, fig.height = 11}

par(mfrow = c(3, 1))
for (sim.name in names(sims)) {
  plot.simulation(sims[[sim.name]], graph.colors, kind = "NS")
}

```

While ```r gs$ptp$name```, ```r gs$ba$name```, and ```r gs$er$name``` networks demonstrate similiar behavior (all three start decreasing rapidly almost immediately after the simulation is started, reaching an inverse plateau in a while), the ```r gs$ws$name``` remains unreceptive to epidemy much longer (up to the second time unit) -- it is a challenge to figure out why.

First, let's take a look at a graph decriptives comparison table:

```{r, results = "asis"}

print.descriptives.for.graphs = function(graphs.list,
                                         include.density = T,
                                         include.diameter = T,
                                         include.clustering.coefficient = T,
                                         include.cluster.count = T) {
  tbl = NULL
  
  for(g.name in names(graphs.list)) {
    g = graphs.list[[g.name]]
    
    tbl.column = matrix()
    colnames(tbl.column) = g$name
    rownames(tbl.column) = c(NA)
    
    if (include.density) {
      g.dens = graph.density(g)
      tbl.column = rbind(tbl.column,
                         g.dens)
      rownames(tbl.column)[length(rownames(tbl.column))] = "Density"
    }
    if (include.diameter) {
      g.diam = diameter(g)
      tbl.column = rbind(tbl.column,
                         g.diam)
      rownames(tbl.column)[length(rownames(tbl.column))] = "Diameter"
    }
    if (include.clustering.coefficient) {
      g.cc = transitivity(g, type = "global")
      tbl.column = rbind(tbl.column,
                         g.cc)
      rownames(tbl.column)[length(rownames(tbl.column))] = "Clustering Coefficient"
    }
    if (include.cluster.count) {
      g.c.no = clusters(g)$no
      tbl.column = rbind(tbl.column,
                         g.c.no)
      rownames(tbl.column)[length(rownames(tbl.column))] = "Cluster Count"
    }
    
    if (is.null(tbl)) {
      tbl = tbl.column
    } else {
      tbl = cbind(tbl,
                  tbl.column)
    }
  }
  
  tbl = tbl[-1, ]
  tbl = xtable(tbl, 
               caption = "Graph Comparison")
  digits(tbl) = c(0, rep(4, length(graphs.list)))
  align(tbl) = paste0("r", 
                      paste0(rep("|c", length(graphs.list)),
                             collapse = ""))
  print(tbl,
        type = "latex",
        floating = T,
        include.rownames = T,
        latex.environments = "center",
        caption.placement = "top")
}


print.descriptives.for.graphs(gs)

```

Clearly, the ```r gs$ws$name```'s diameter is outstanding, which justifies the long-running SIR simuluation: It is just the epidemy that needs a fair amount of time to spread.



\newpage
### NI Comparison Charts

```{r, fig.width = 10, fig.height = 11}

par(mfrow = c(3, 1))
for (sim.name in names(sims)) {
  plot.simulation(sims[[sim.name]], graph.colors, kind = "NI")
}

```

Except for the ```r gs$ws$name``` series, other types of network expose similiar behaviour over time, each hitting its peak around the first time unit. By contrast, due to the ```r gs$ws$name```'s diameter infection spread is negligible (even for $\beta \gg \gamma$, its footprint is nowhere close to that in other models).



\newpage
### NR Comparison Charts

```{r, fig.width = 10, fig.height = 11}

par(mfrow = c(3, 1))
for (sim.name in names(sims)) {
  plot.simulation(sims[[sim.name]], graph.colors, kind = "NR")
}

```

Once again, the good-old ```r gs$ws$name``` stands out: for $\beta \gg \gamma$, the number of recovered actors is nearly 0; for $\beta = \gamma$, a small number of previously infected actors recover in a graduate manner; the last plot exposes a slight ununiformity in NR by the very end of simulation, which is likely justified by the respective networks structure.
