---
title: "Social Network Analysis <br/> Home Assignment 2"
author: 'Shupeyko Nikita'
date: "due date - 15.05.2016 23:59"
output:
  pdf_document:
    latex_engine: xelatex
    toc: yes
  html_document:
    toc: yes
---
# Graph models. Centrality metrics

```{r, echo = F, message = F, warning = F}

library(igraph)
library(rgexf)
library(xtable)
library(R.matlab)

```



## Task 1. Your social network

For the first task, you have to load your [vk.com](http://vk.com) network. Please follow the instructions posted on the course wiki or user [VK Application](vk.com/app3861133). For [FB](http://facebook.com) users try to use [NetVizz](http://thepoliticsofsystems.net/2010/03/netvizz-facebook-to-gephi/). If you did it correctly, you should have a GraphML file with your own network. Read it to R:

```{r, warning = F, fig.width = 10, fig.height= 10}

# Social graphs ("Social Graph") acquired through 
# [Export friends graph](vk.com/app3861133) VK app.

# Friends graph relative path
# (note the file is not available on remote due to privacy issues -- 
# feel free to generate one on your own):
g_vkf.rel.path = "vk-friends-20049226.gexf"
# Friends-and-followers graph relative path:
g_vkff.rel.path = ""

g.rel.path = g_vkf.rel.path

g = gexf.to.igraph(read.gexf(g.rel.path))

g.n = vcount(g)
g.m = ecount(g)
g.avg.deg = 2*g.m / g.n

g.v.names = names(V(g))

# 'Anonimizing' g:
g.adj.mat = get.adjacency(g)
g.adj.mat.names = 1:g.n
colnames(g.adj.mat) = g.adj.mat.names
rownames(g.adj.mat) = colnames(g.adj.mat)
g = graph.adjacency(g.adj.mat, add.colnames = T)

g.deg = degree(g)

```



### 1. Degree distribution

First, plot degree distribution of your network in log-log scales:

```{r, warning = F, fig.width = 10, fig.height = 7}

# For further power-law fitting correctness, let degree n <-> degree n+1:
g.deg.plus1 = g.deg + 1

# Obtaining empirical degree distribution:
g.deg.plus1.rng.min = min(g.deg.plus1)
g.deg.plus1.rng.max = max(g.deg.plus1)
g.deg.plus1.rng = as.numeric(g.deg.plus1.rng.min:g.deg.plus1.rng.max)
g.deg.plus1.dist = degree.distribution(g)[g.deg.plus1.rng]

# Constructing theoretical power-law degree distribution:
g.pl.fit = power.law.fit(g.deg.plus1, implementation = "plfit")
func_pl = function(x) 
    return ((g.pl.fit$alpha-1) * g.pl.fit$xmin^(g.pl.fit$alpha-1) * x^(-g.pl.fit$alpha))

# For the sake of clarity, the empirical and theoretical 
# degree distributions are plotted side-by-side
plot(g.deg.plus1.dist, log = "xy", 
     main = "Social Graph Node Degree Distribution", xlab = "Node degree (log)",
     ylab = "Degree distribution (log)")
par(new = T)
curve(func_pl, from = g.deg.plus1.rng.min, to = g.deg.plus1.rng.max, log="xy", 
      main = "", xlab = "", ylab = "", col = "red", axes = F, add = F)
par(new = F)
legend("bottomleft", 
       legend = c("Empirical", sprintf("Theoretical power-law fit (alpha = %.4f)", g.pl.fit$alpha)), 
       col = c('black', 'red'), lty = 1)

```

Is there any correspondence between actual degree distribution of your network and the Power Law distribution? If not, explain why.

*According to KS test results, **Power Law fits Social Graph** node degree distribution **with the probability of `r sprintf("%.4f", g.pl.fit$KS.p)`**.*

Now, let's see how it would look if it was random. Produce Erdos-Renyi graph matching your real network (same number of nodes and same average degree). Compare it with your degree distribution.

```{r, warning = F, fig.width = 10, fig.height = 5}

# Since both the original and the produced graph are required to have the same number
# of nodes and average node degree, the number of Erdos-Renyi graph edges, by necessity,
# equals the latter in the original graph.
g_erg = sample_gnm(n = g.n, m = g.m, directed = T)

g_erg.n = vcount(g_erg)
g_erg.m = ecount(g_erg)

g_erg.deg.dist = degree.distribution(g_erg)

plot(g.deg.plus1.dist, log = "xy", 
     main = "Social Graph vs. Erdos-Renyi", xlab = "Node degree (log)", 
     ylab = "Degree distribution (log)")
par(new = T)
plot(g_erg.deg.dist, log="xy",
     main = "", xlab = "", ylab = "", col="red", axes = F)
par(new = F)
legend("bottomleft", legend = c("Empirical", "Erdos-Renyi"), 
       col = c('black', 'red'), lty = 1)

```



### 2. Compute centrality metrics

Compute for your network: 

* degree centrality

* closeness centrality

* betweenness centrality

* eigenvector centrality

* Bonacich power centrality

* Alpha centrality

```{r}

g.deg.cent = g.deg
g.cls.cent = closeness(g)
g.btw.cent = betweenness(g)
g.ev.cent = eigen_centrality(g)
g.bp.cent = bonpow(g)
g.alpha.cent = alpha.centrality(g)

```

Output six plots corresponding to six centrality metrics you've computed: 

* Use first names of your friends as node labels on the graph (you may hide this information if you wish -- change it by integer ID).
* Keep the same layout of the network.
* Make node sizes and colours proportional to the respective centrality metrics.

```{r, fig.width = 10, fig.height = 10}

g.cent.th = 100
g.cent.palette = colorRampPalette(c("blue", "red"))
g.cent.lyt = layout.fruchterman.reingold(g)
g.cent.vclr = g.cent.palette(g.cent.th)[as.numeric(cut(g.deg.cent, breaks = g.cent.th))]

func_scale01 = function(l) scale(l, center = min(l), scale = max(l) - min(l))
func_cent.vert.size = function(l) 0 + 5 * func_scale01(l)

plot(g, layout = g.cent.lyt, 
     main = "Social Graph Degree Centrality",
     vertex.color = g.cent.vclr, vertex.size = func_cent.vert.size(g.deg.cent))

plot(g, layout = g.cent.lyt, 
     main = "Social Graph Closeness Centrality", 
     vertex.color = g.cent.vclr, vertex.size = func_cent.vert.size(g.cls.cent))

plot(g, layout = g.cent.lyt, 
     main = "Social Graph Betweeness Centrality",
     vertex.color = g.cent.vclr, vertex.size = func_cent.vert.size(g.btw.cent))

plot(g, layout = g.cent.lyt, 
     main = "Social Graph Eigenvector Centrality",
     vertex.color = g.cent.vclr, vertex.size = func_cent.vert.size(g.ev.cent$vector))

plot(g, layout = g.cent.lyt,
     main = "Social Graph Bonachich Centrality", 
     vertex.color = g.cent.vclr, vertex.size = func_cent.vert.size(g.bp.cent))

plot(g, layout = g.cent.lyt, main = "Social Graph Alpha Centrality", 
     vertex.color = g.cent.vclr, vertex.size = func_cent.vert.size(g.alpha.cent))

```

Now, output top ten nodes in each ranking. Again, print only first names in your table to keep privacy:

```{r results = "asis"}

func_topN = function(n, l) which( l > sort(l)[ length(l)-n ] )
func_top10 = function(l) func_topN(10, l)
func_before.space = function(s) sub("\\s.*", "", s)


options(xtable.comment = F)

func_print.custom.xtable = function(m) 
    print(xtable(m), floating = F, type = "latex", include.rownames = F)

g.deg.cent.top10 = func_top10(g.deg.cent)
g.deg.cent.top10.mat = cbind(func_before.space( g.v.names[g.deg.cent.top10] ), g.deg.cent.top10)
colnames(g.deg.cent.top10.mat) = c("Node", "Degree Centrality")
func_print.custom.xtable(g.deg.cent.top10.mat)

g.cls.cent.top10 = func_top10(g.cls.cent)
g.cls.cent.top10.mat = cbind(func_before.space( g.v.names[g.cls.cent.top10] ), g.cls.cent.top10)
colnames(g.cls.cent.top10.mat) = c("Node", "Closeness Centrality")
func_print.custom.xtable(g.cls.cent.top10.mat)

g.btw.cent.top10 = func_top10(g.btw.cent)
g.btw.cent.top10.mat = cbind(func_before.space(g.v.names[g.btw.cent.top10]), g.btw.cent.top10)
colnames(g.btw.cent.top10.mat) = c("Nodes", "Betweeness Centrality")
func_print.custom.xtable(g.btw.cent.top10.mat)

g.ev.cent.top10 = func_top10(g.ev.cent$vector)
g.ev.cent.top10.mat = cbind(func_before.space(g.v.names[g.ev.cent.top10]), g.ev.cent.top10)
colnames(g.ev.cent.top10.mat) = c("Nodes", "Eigenvector Centrality")
func_print.custom.xtable(g.ev.cent.top10.mat)

g.bp.cent.top10 = func_top10(g.bp.cent)
g.bp.cent.top10.mat = cbind(func_before.space(g.v.names[g.bp.cent.top10]), g.bp.cent.top10)
colnames(g.bp.cent.top10.mat) = c("Nodes", "Bonachich Centrality")
func_print.custom.xtable(g.bp.cent.top10.mat)

g.alpha.cent.top10 = func_top10(g.alpha.cent)
g.alpha.cent.top10.m = cbind(func_before.space(g.v.names[g.alpha.cent.top10]), g.alpha.cent.top10)
colnames(g.alpha.cent.top10.m) = c("Nodes", "Alpha Centrality")
func_print.custom.xtable(g.alpha.cent.top10.m)

```

*Daniil has predictably revealed the highest betweeness centrality in my social graph since he is one of my most communicative buddies, who indeed preserves strong ties within our group.*



## Task 2. Flickr network

Please download [flickr.mat](http://www.hse.ru/data/2015/05/01/1098247366/flickr.mat) 

Data contains sparse matrix A and list of user names. This is a denser part of the Flickr photo sharing site friendship graph from 2006. Edge direction corresponds to friendship requests (following). Some of the links are reciprocal, others not.

It's a Matlab file. How to deal with it in R? There is a package `R.matlab`. Please install it and call `library(R.matlab)`

Now use `readMat` function to read the file and extract adjacency matrix and a list of user names:

```{r, warning = F}

flickr = readMat("flickr.mat")
flickr.mat = as.matrix(flickr[1]$A)
flickr.names = flickr[2]$names

```

Look at user names. You might want to remove spaces from the names. Use a function `gsub` to remove them:

```{r}

flickr.names = gsub(" ", "", flickr.names)

```

Now create a graph, output the number of vertices and edges:

```{r}

f_g = graph.adjacency(flickr.mat)

f_g.n = vcount(f_g)
f_g.m = ecount(f_g)

f_g.v.names = flickr.names

paste0("Vertex count: ", f_g.n)
paste0("Edge count: ", f_g.m)

```

Compute in- and out- degree centralities, PageRank, Hubs and Authorities for this network:

```{r}

f_g.indeg.cent = degree(f_g, mode = "in")
f_g.outdeg.cent = degree(f_g, mode = "out")
f_g.pr = page.rank(f_g)
f_g.hub = hub.score(f_g)
f_g.auth = authority.score(f_g)

```

Print top ten names in each ranking:

```{r results = "asis"}

f_g.indeg.cent.top10 = func_top10(f_g.indeg.cent)
f_g.indeg.cent.top10.m = cbind(f_g.v.names[f_g.indeg.cent.top10], f_g.indeg.cent.top10)
colnames(f_g.indeg.cent.top10.m) = c("Node", "In-degree Centrality")
func_print.custom.xtable(f_g.indeg.cent.top10.m)

f_g.outdeg.cent.top10 = func_top10(f_g.v.names)
f_g.outdeg.cent.top10.mat = cbind(f_g.v.names[f_g.outdeg.cent.top10], f_g.outdeg.cent.top10)
colnames(f_g.outdeg.cent.top10.mat) = c("Node", "Out-degree Centrality")
func_print.custom.xtable(f_g.outdeg.cent.top10.mat)

f_g.pr.top10 = func_top10(f_g.pr$vector)
f_g.pr.top10.mat = cbind(f_g.v.names[f_g.pr.top10], f_g.pr.top10)
colnames(f_g.pr.top10.mat) = c("Node", "Pagerank")
func_print.custom.xtable(f_g.pr.top10.mat)

f_g.hub.top10 = func_top10(f_g.hub$vector)
f_g.hub.top10.mat = cbind(f_g.v.names[f_g.hub.top10], f_g.hub.top10)
colnames(f_g.hub.top10.mat) = c("Node", "Hub")
func_print.custom.xtable(f_g.hub.top10.mat)

f_g.auth.top10 = func_top10(f_g.auth$vector)
f_g.auth.top10.mat = cbind(f_g.v.names[f_g.auth.top10], f_g.auth.top10)
colnames(f_g.auth.top10.mat) = c("Node", "Authority")
func_print.custom.xtable(f_g.auth.top10.mat)

```

Produce the following plots:

* In-degree centralities versus out-degree centralities

* In-degree centralities versus authorities

* Out-degree centralities versus hubs

* Hubs versus authorities

* PageRank versus hubs

* PageRank versus authorities


```{r, fig.height = 7, fig.width=10}

func_plot.vs = function(m1, lab1, m2, lab2, col = "blue") 
    plot(m1, m2, xlab = lab1, ylab = lab2, col = col)

par(mfrow = c(2, 3))

func_plot.vs(f_g.indeg.cent, "Indegree", f_g.outdeg.cent, "Out-degree")

func_plot.vs(f_g.indeg.cent, "Indegree", f_g.auth$vector, "Authority")

func_plot.vs(f_g.outdeg.cent, "Out-degree", f_g.hub$vector, "Hub")

func_plot.vs(f_g.hub$vector, "Hub", f_g.auth$vector, "Authority")

func_plot.vs(f_g.pr$vector, "Pagerank", f_g.hub$vector, "Hub")

func_plot.vs(f_g.pr$vector, "Pagerank", f_g.auth$vector, "Authority")

mtext("Centrality Comparison", outer = T)

```

*For \<in-degree, authority\>, \<out-degree, hub\> and \<pagerank, authority\> pairs a positive correlation is revealed, as opposed to the rest of those being compared.*

